---
title: "Kalman, Time series and Markov Chain"
author: "RR"
date: "12 luglio 2016"
output: html_document
---


##Kalman
For more details on Kalman Filter see <https://cran.r-project.org/web/packages/FKF/FKF.pdf>.

Theory and explanation are done in this work <https://www.jstatsoft.org/article/view/v039i02/v39i02.pdf>

###First Example

```{r}
library("FKF",lib.loc="~/R/win-library/3.2")
## <--------------------------------------------------------------------------->
## Example 1: ARMA(2, 1) model estimation.
## <--------------------------------------------------------------------------->
## This example shows how to fit an ARMA(2, 1) model using this Kalman
## filter implementation (see also stats' makeARIMA and KalmanRun).
n <- 1000
## Set the AR parameters
ar1 <- 0.6
ar2 <- 0.2
ma1 <- -0.2
sigma <- sqrt(0.2)
## Sample from an ARMA(2, 1) process
a <- arima.sim(model = list(ar = c(ar1, ar2), ma = ma1), n = n,
innov = rnorm(n) * sigma)
## Create a state space representation out of the four ARMA parameters
arma21ss <- function(ar1, ar2, ma1, sigma) {
Tt <- matrix(c(ar1, ar2, 1, 0), ncol = 2)
Zt <- matrix(c(1, 0), ncol = 2)
ct <- matrix(0)
dt <- matrix(0, nrow = 2)
GGt <- matrix(0)
H <- matrix(c(1, ma1), nrow = 2) * sigma
HHt <- H %*% t(H)
a0 <- c(0, 0)
P0 <- matrix(1e6, nrow = 2, ncol = 2)
return(list(a0 = a0, P0 = P0, ct = ct, dt = dt, Zt = Zt, Tt = Tt, GGt = GGt, HHt = HHt))
}
## The objective function passed to 'optim'
objective <- function(theta, yt) {
sp <- arma21ss(theta["ar1"], theta["ar2"], theta["ma1"], theta["sigma"])
ans <- fkf(a0 = sp$a0, P0 = sp$P0, dt = sp$dt, ct = sp$ct, Tt = sp$Tt,
Zt = sp$Zt, HHt = sp$HHt, GGt = sp$GGt, yt = yt)
return(-ans$logLik)
}
theta <- c(ar = c(0, 0), ma1 = 0, sigma = 1)
fit <- optim(theta, objective, yt = rbind(a), hessian = TRUE)
fit
## Confidence intervals
rbind(fit$par - qnorm(0.975) * sqrt(diag(solve(fit$hessian))),
fit$par + qnorm(0.975) * sqrt(diag(solve(fit$hessian))))
## Filter the series with estimated parameter values
sp <- arma21ss(fit$par["ar1"], fit$par["ar2"], fit$par["ma1"], fit$par["sigma"])
ans <- fkf(a0 = sp$a0, P0 = sp$P0, dt = sp$dt, ct = sp$ct, Tt = sp$Tt,
Zt = sp$Zt, HHt = sp$HHt, GGt = sp$GGt, yt = rbind(a))
## Compare the prediction with the realization
plot(ans, at.idx = 1, att.idx = NA, CI = NA)
lines(a, lty = "dotted")
## Compare the filtered series with the realization
plot(ans, at.idx = NA, att.idx = 1, CI = NA)
lines(a, lty = "dotted")
## Check whether the residuals are Gaussian
plot(ans, type = "resid.qq")
## Check for linear serial dependence through 'acf'
plot(ans, type = "acf")
```           


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

###Second Example

```{r}
library("FKF",lib.loc="~/R/win-library/3.2")
## <--------------------------------------------------------------------------->
## Example 2: Local level model for the Nile's annual flow.
## <--------------------------------------------------------------------------->
## Transition equation:
## alpha[t+1] = alpha[t] + eta[t], eta[t] ~ N(0, HHt)
## Measurement equation:
## y[t] = alpha[t] + eps[t], eps[t] ~ N(0, GGt)
y <- Nile
y[c(3, 10)] <- NA # NA values can be handled
## Set constant parameters:
dt <- ct <- matrix(0)
Zt <- Tt <- matrix(1)
a0 <- y[1] # Estimation of the first year flow
P0 <- matrix(100) # Variance of 'a0'
## Estimate parameters:
fit.fkf <- optim(c(HHt = var(y, na.rm = TRUE) * .5,
GGt = var(y, na.rm = TRUE) * .5),
fn = function(par, ...)
-fkf(HHt = matrix(par[1]), GGt = matrix(par[2]), ...)$logLik,
yt = rbind(y), a0 = a0, P0 = P0, dt = dt, ct = ct,
Zt = Zt, Tt = Tt, check.input = FALSE)
## Filter Nile data with estimated parameters:
fkf.obj <- fkf(a0, P0, dt, ct, Tt, Zt, HHt = matrix(fit.fkf$par[1]),
GGt = matrix(fit.fkf$par[2]), yt = rbind(y))
## Compare with the stats' structural time series implementation:
fit.stats <- StructTS(y, type = "level")
fit.fkf$par
fit.stats$coef
## Plot the flow data together with fitted local levels:
plot(y, main = "Nile flow")
lines(fitted(fit.stats), col = "green")
lines(ts(fkf.obj$att[1, ], start = start(y), frequency = frequency(y)), col = "blue")
legend("top", c("Nile flow data", "Local level (StructTS)", "Local level (fkf)"),
col = c("black", "green", "blue"), lty = 1)


```


###Third Example

```{r}
library("FKF",lib.loc="~/R/win-library/3.2")
## Local level model for the treering width data.
## Transition equation:
## alpha[t+1] = alpha[t] + eta[t], eta[t] ~ N(0, HHt)
## Measurement equation:
## y[t] = alpha[t] + eps[t], eps[t] ~ N(0, GGt)
y <- treering
y[c(3, 10)] <- NA # NA values can be handled
## Set constant parameters:
dt <- ct <- matrix(0)
Zt <- Tt <- matrix(1)
a0 <- y[1] # Estimation of the first width
P0 <- matrix(100) # Variance of 'a0'
## Estimate parameters:
fit.fkf <- optim(c(HHt = var(y, na.rm = TRUE) * .5,
GGt = var(y, na.rm = TRUE) * .5),
fn = function(par, ...)
-fkf(HHt = matrix(par[1]), GGt = matrix(par[2]), ...)$logLik,
yt = rbind(y), a0 = a0, P0 = P0, dt = dt, ct = ct,
Zt = Zt, Tt = Tt, check.input = FALSE)
## Filter Nile data with estimated parameters:
fkf.obj <- fkf(a0, P0, dt, ct, Tt, Zt, HHt = matrix(fit.fkf$par[1]),
GGt = matrix(fit.fkf$par[2]), yt = rbind(y))
## Plot the width together with fitted local levels:
plot(y, main = "Treering data")
lines(ts(fkf.obj$att[1, ], start = start(y), frequency = frequency(y)), col = "blue")
legend("top", c("Treering data", "Local level"), col = c("black", "blue"), lty = 1)
## Check the residuals for normality:
plot(fkf.obj, type = "resid.qq")
## Test for autocorrelation:
plot(fkf.obj, type = "acf", na.action = na.pass)
```

##Time Series
<https://cran.r-project.org/web/packages/timeSeries/timeSeries.pdf>
```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data Set -
x <- MSFT
## Aggregate by Weeks -
by <- timeSequence(from = start(x), to = end(x), by = "week")
aggregate(x, by, mean)
## Aggregate to Last Friday of Month -
by <- unique(timeLastNdayInMonth(time(x), 5))
X <- aggregate(x, by, mean)
X
dayOfWeek(time(X))
isMonthly(X)
## Aggregate to Last Day of Quarter -
by <- unique(timeLastDayInQuarter(time(x)))
X <- aggregate(x, by, mean)
X
isQuarterly(X)
## Aggregate daily records to end of month records -
X <- daily2monthly(x)
X
isMonthly(X)
## Aggregate da, ily records to end of week records -
X <- daily2weekly(x, startOn="Fri")
X
dayOfWeek(time(X))
plot(x)

```


###First Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Use MSFT and Compute Sample Size -
dim(MSFT)
## Align the Series -
MSFT.AL <- align(MSFT)
## Show the Size of the Aligned Series -
dim(MSFT.AL)
```

###Second Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Percentual Returns of Swiss Bond Index and Performance Index -
LPP <- 100 * LPP2005REC[, c("SBI", "SPI")]
head(LPP, 20)
## Aggregate Quarterly Returns -
applySeries(LPP, by = "quarterly", FUN = colSums)
## Aggregate Quarterly every last Friday in Quarter -
oneDay <- 24*3600
from <- unique(timeFirstDayInQuarter(time(LPP))) - oneDay
from <- timeLastNdayInMonth(from, nday = 5)
to <- unique(timeLastDayInQuarter(time(LPP)))
to <- timeLastNdayInMonth(to, nday = 5)
data.frame(from = as.character(from), to = as.character(to))
applySeries(LPP, from, to, FUN = colSums)
## Count Trading Days per Month -
colCounts <- function(x) rep(NROW(x), times = NCOL(x))
applySeries(LPP, FUN = colCounts, by = "monthly")
## Alternative Use -
fapply(LPP, from, to, FUN = colSums)
```

###Third Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Create an Artificial timeSeries Object -
setRmetricsOptions(myFinCenter = "GMT")
charvec <- timeCalendar()
data <- matrix(rnorm(12))
TS <- timeSeries(data, charvec, units = "RAND")
TS
## Coerce to Vector -
as.vector(TS)
## Coerce to Matrix -
as.matrix(TS)
## Coerce to Data Frame -
as.data.frame(TS)
```

###Fourth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data Set -
x <- MSFT[1:10, ]
colnames(x)
## Attach the Series and Compute the Range -
attach(x)
range <- High - Low
range
## Convert Vector to a timeSeries Object -
timeSeries(data=range, charvec=time(x), units="Range")
## Detach the series from the search path -

ans <- try(High, silent=TRUE)
cat(ans[1])
detach("x")
```

###Fifth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Create an artificial timeSeries Object -
tS <- dummySeries()
tS
## Get Optional Attributes -
getAttributes(tS)
tS@documentation
## Set a new Optional Attribute -
setAttributes(tS) <- list(what="A dummy Series")
tS
getAttributes(tS)
tS@documentation
plot(ts)
```

###Sixth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data Set -
x <- MSFT[1:12, ]
x
## Bind Columnwise -
X <- cbind(x[, "Open"], returns(x[, "Open"]))
colnames(X) <- c("Open", "Return")
X
## Bind Rowwise -
Y <- rbind(x[1:3, "Open"], x[10:12, "Open"])
Y
plot(x)
plot(y)
```

###Seventh Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Simulated Return Data -
x = matrix(rnorm(24), ncol = 2)
## Cumulative Sums Column by Column -
colCumsums(x)
```

###Eighth Example

```{r}
## Simulated Return Data in Matrix Form -
x = matrix(rnorm(252), ncol = 2)
## Mean Columnwise Statistics -
colStats(x, FUN = mean)
## Quantiles Column by Column -
colQuantiles(x, prob = 0.10, type = 1)
```

###Nineth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Get Description from timeSeries -
comment(LPP2005REC)
## Add User to comment -
comment(LPP2005REC) <- paste(comment(LPP2005REC), "by User Rmetrics")
comment(LPP2005REC)
```

###Tenth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Use the Microsofts' Close Prices Indexed to 1 -
MSFT.CL <- MSFT[, "Close"]
MSFT.CL <- MSFT.CL/MSFT[[1, "Close"]]
head(MSFT.CL)
## Compute Discrete Return -
MSFT.RET <- returns(MSFT.CL, method = "discrete")
## Cumulated Series and Compare -
MSFT.CUM <- cumulated(MSFT.RET, method = "discrete")
head(cbind(MSFT.CL, MSFT.CUM))

```

###Eleventh Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data -
X <- MSFT[1:10, 1:4]
## Get Data Part -
DATA <- getDataPart(X)
class(DATA)
```

###Twelveth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data Set -
x <- MSFT[1:12, ]
x
## Compute Differences -
diff(x)
## Trimmed Differences -
diff(x, trim=TRUE)
## Padded Differences -
diff(x, trim=FALSE, pad=0)
```

###Thirteenth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Swiss Pension Fund Benchmark Data -
X <- LPP2005REC[1:10, 1:3]
## Get Dimension -
dim(X)
## Get Column and Row Names -
dimnames(X)
## Get Column / Row Names -
colnames(X)
rownames(X)
## Try your own DIM -
DIM <- function(x) {c(NROW(x), NCOL(x))}
DIM(X)
DIM(X[, 1])
## Try length / LENGTH -
length(X)
length(X[, 1])
LENGTH <- function(X) NROW(X)
LENGTH(X)
## Columns / Rows -
ncol(X); NCOL(X)
nrow(X); NROW(X)
## See also -
isUnivariate(X)
isMultivariate(X)
```
###Fourteenth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Use Swiss Pension Fund Data Set of Returns -
head(LPP2005REC)
SPI <- LPP2005REC[, "SPI"]
head(SPI)
## Plot Drawdowns -
dd = drawdowns(LPP2005REC[, "SPI"], main = "Drawdowns")
plot(dd)
dd = drawdowns(LPP2005REC[, 1:6], main = "Drawdowns")
plot(dd)
## Compute Drawdowns Statistics -
ddStats <- drawdownsStats(SPI)
class(ddStats)
ddStats
## Note, Only Univariate Series are allowd -
ddStats <- try(drawdownsStats(LPP2005REC))
class(ddStats)

```
###Fifteenth Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Compute Durations in days for the MSFT Sereries -
head(durations(MSFT, units = "days"))
head(durations(MSFT, trim = TRUE, units = "days"))
## The same in hours -
head(durations(MSFT, trim = TRUE, units = "hours"))

```
###Linear filtering Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Creata a Dummy Signal 'timeSeries' -
data <- matrix(rnorm(100), ncol = 2)
s <- timeSeries(data, units=c("A", "B"))
head(s)
## Filter the series -
f <- filter(s, rep(1, 3))
head(f)
## Plot and Compare the first series -
plot(cbind(s[, 1], f[, 1]), plot.type="s")
```
###Financial Center Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## An artificial timeSeries Object -
tS <- dummySeries()
tS
## Print Financial Center -
finCenter(tS)
getFinCenter(tS)
## Assign New Financial Center -
finCenter(tS) <- "Zurich"
tS
setFinCenter(tS) <- "New_York"
tS
```
###timeSeries Class, Coercion and Transformation Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Create an Artificial timeSeries Object -
setRmetricsOptions(myFinCenter = "GMT")
charvec <- timeCalendar()
data <- matrix(rnorm(12))
TS <- timeSeries(data, charvec, units = "RAND")
TS
## Test for timeSeries -
is.timeSeries(TS)

```
###CChecks if a Time Series is Univariate Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data -
setRmetricsOptions(myFinCenter = "GMT")
data(MSFT)
Open = MSFT[, "Open"]
## Is the timeSeries Univariate -
isUnivariate(MSFT)
isUnivariate(Open)
## Is the timeSeries Multivariate -
isMultivariate(MSFT)
isMultivariate(Open)
```
###Mathematical Time Series Operations Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Create an Artificial timeSeries Object -
setRmetricsOptions(myFinCenter = "GMT")
charvec = timeCalendar()
set.seed(4711)
data = matrix(exp(cumsum(rnorm(12, sd = 0.1))))
TS = timeSeries(data, charvec, units = "TS")
TS
## Mathematical Operations: | +/- * ^ ... -
TS^2
TS[2:4]
OR = returns(TS)
OR
OR > 0
```
###Lag a Time Series Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Micsrosoft Data Set -
x = MSFT[1:20, "Open"]
## Lag the timeSeries Object:
lag(x, k = -1:1)

```
###Merges two ’timeSeries’ objects Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Series -
x <- MSFT[1:12, ]
## Merge 'timeSeries' with missing Object -
merge(x)
## Merge 'timeSeries' with numeric Object -
y <- rnorm(12)
class(y)
merge(x, y)
## Merge 'timeSeries' with matrix Object -
y <- matrix(rnorm(24), ncol=2)
class(y)
merge(x, y)
## Merge 'timeSeries' with matrix Object -
y <- timeSeries(data=rnorm(12), charvec=time(x))
class(y)
merge(x, y)
```
###Model Frames for Time Series Objects  Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data -
setRmetricsOptions(myFinCenter = "GMT")
X <- MSFT[1:12, ]
## Extract High's and Low's:
DATA <- model.frame( ~ High + Low, data = X)
class(DATA)
as.timeSeries(DATA)
## Extract Open Prices and their log10's:
base <- 10
Open <- model.frame(Open ~ log(Open, base = `base`), data = X)
colnames(Open) <- c("X", "log10(X)")
class(Open)
as.timeSeries(Open)
```
###Plot a time-serie Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Swiss Pension Fund Benchmark Data -
LPP <- LPP2005REC[1:12, 1:4]
colnames(LPP) <- abbreviate(colnames(LPP), 2)
finCenter(LPP) <- "GMT"
## Example Plot 1 -
plot(LPP[, 1], type = "o", col = "steelblue",
main = "LPP", xlab = "2005", ylab = "Return")
plot(LPP[, 1], at="auto", type = "o", col = "steelblue",
main = "LPP", xlab = "2005", ylab = "Return")
## Example Plot 2 -
plot(LPP[, 1:2], type = "o", col = "steelblue",
main = "LPP", xlab = "2005", ylab = "Return")
## Example Plot 3 -
plot(LPP[, 1], LPP[, 2], type = "p", col = "steelblue",
main = "LPP", xlab = "Return 1", ylab = "Return 2")
## Example Plot 4a, The Wrong Way to do it! -
LPP <- as.timeSeries(data(LPP2005REC))
ZRH <- as.timeSeries(LPP[,"SPI"], zone = "Zurich", FinCenter = "Zurich")
NYC <- as.timeSeries(LPP[,"LMI"], zone = "NewYork", FinCenter = "NewYork")
finCenter(ZRH)
finCenter(NYC)
plot(ZRH, at="auto", type = "p", pch = 19, col = "blue")
points(NYC, pch = 19, col = "red")
## Example Plot 4b, Convert NYC to Zurich Time -
finCenter(ZRH) <- "Zurich"
finCenter(NYC) <- "Zurich"
at <- unique(round(time(ZRH)))
plot(ZRH, type = "p", pch = 19, col = "blue", format = "%b %d", at = at,
xlab = paste(ZRH@FinCenter, "local Time"), main = ZRH@FinCenter)
points(NYC, pch = 19, col = "red")
## Example 4c, Force Everything to GMT Using "FinCenter" Argument -
finCenter(ZRH) <- "Zurich"
finCenter(NYC) <- "NewYork"
at <- unique(round(time(ZRH)))
plot(ZRH, type = "p", pch = 19, col = "blue", format = "%b %d", at = at,
FinCenter = "GMT", xlab = "GMT", main = "ZRH - GMT")
points(NYC, FinCenter = "GMT", pch = 19, col = "red")

```
###Print a Time Series Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Micsrosoft Data -
setRmetricsOptions(myFinCenter = "GMT")
LPP <- MSFT[1:12, 1:4]
## Abbreviate Column Names -
colnames(LPP) <- abbreviate(colnames(LPP), 6)
## Print Data Set -
print(LPP)
## Alternative Use, Show Data Set -
show(LPP)

```
###Sample Ranks of a Time Series Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load Microsoft Data -
X <- 100 * returns(MSFT)
## Compute the Ranks -
head(rank(X[, "Open"]), 10)
## Only Interested in the Vector, then use -
head(rank(series(X[, "Open"])), 10)

```
###Rolling Statistics Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Use Swiss Pension Fund Data Set of Returns -
head(LPP2005REC)
SPI <- LPP2005REC[, "SPI"]
head(SPI)
## Plot Drawdowns -
rmean <- rollMean(SPI, k = 10)
plot(rmean)
```
###Cumulated Column Statistics Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Simulated Monthly Return Data -
X = matrix(rnorm(24), ncol = 2)
## Compute cumulated Sums -
rowCumsums(X)
```
###Get and Set Data of a ’timeSeries’ Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## A Dummy timeSeries Object
ts <- timeSeries()
ts
## Get the Matrix Part -
mat <- series(ts)
class(mat)
mat
## Assign a New Univariate Series -
series(ts) <- rnorm(12)
ts
## Assign a New Bivariate Series -
series(ts) <- rnorm(12)
ts
plot(ts)
```
###extract window Example

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
## Load LPP Benchmark Returns -
x <- LPP2005REC[, 7:9]
range(time(x))
## Extract Data for January 2006 -
window(x, "2006-01-01", "2006-01-31")
```
###More plotting Example
Taken from <https://cran.r-project.org/web/packages/timeSeries/vignettes/timeSeriesPlot.pdf>

```{r}
library("timeSeries", lib.loc="~/R/win-library/3.2")
library(PerformanceAnalytics)
require(timeSeries)
require(xts)
require(fTrading)
tS1 <- 100 * cumulated(LPP2005REC[, 1]) # SBI (univariate)
tS2 <- 100 * cumulated(LPP2005REC[, 1:2]) # SBI & SPI (bivariate)
tS3 <- 100 * cumulated(LPP2005REC[, 1:3]) # SBI, SPI, SWIIT (Swiss Market)
tS6 <- 100 * cumulated(LPP2005REC[, 1:6]) # Swiss and Foreign Market Indexes

par(mfrow=c(1, 1))
plot(tS1)

par(mfrow=c(3, 1))
xts::plot.xts(as.xts(tS1))
PerformanceAnalytics::chart.TimeSeries(as.xts(tS1))
plot(tS1)

par(mfrow=c(1, 1))
plot(tS3, plot.type="s")

par(mfrow=c(2, 1))
require(PerformanceAnalytics)
PerformanceAnalytics::chart.TimeSeries(as.xts(tS3))
plot(tS3, plot.type="s")

par(mfrow=c(1, 1))
plot(tS3, plot.type="m")

par(mfrow=c(1, 1))
plot(tS6, plot.type="m")

par(mfrow=c(1, 1))
plot(tS3, plot.type="m", mar=c(gap=0.3, 5.1, gap=0.3, 2.1))

par(mfrow=c(2, 1))
par(mar = c(bottom=1.5, 5.1, top=4, 2.1))
plot(tS2[, 1])
par(mar = c(bottom=4, 5.1, top=1.5, 2.1))
plot(tS2[, 2])

nf <- layout(mat=matrix(c(1, 1, 2, 3), byrow = TRUE, nrow=2))
par(mar = c(bottom=2, 5.1, top=3, 2.1))
plot(tS3[, 1])
par(mar = c(bottom=3, 5.1, top=2, 1.1))
plot(tS3[, 2])
par(mar = c(bottom=3, 4.1, top=2, 2.1))
plot(tS3[, 3])

nf <- layout(mat=matrix(c(1, 1, 2, 3), byrow=TRUE, nrow=2), heights=c(2.5,1))
par(mar = c(bottom=2, 5.1, top=3, 2.1))
plot(tS3[, 1])
par(mar = c(bottom=3, 5.1, top=1.5, 1.1))
plot(tS3[, 2])
par(mar = c(bottom=3, 4.1, top=1.5, 2.1))
plot(tS3[, 3])

par(mfrow=c(1, 1))
lines2 <- function(X, Y, type, xlab, ylab, col, pch, lty, lwd, cex) {lines(x=X, y=Y, col=col)+ abline(h=0, col = "brown", lwd=2)}
plot(returns(tS3), plot.type="m", col = .colorwheelPalette(3), panel=lines2, at="pretty")


par(mfrow=c(1, 1))
lines2 <- function(X, Y, type, xlab, ylab, col, pch, lty, lwd, cex) {lines(x=X, y=Y, type="h", col=col)+ rug(Y, side=4, col="steelblue") }
plot(returns(tS6), plot.type="m", col = .colorwheelPalette(6), panel=lines2, at="pretty")

library("fTrading")
par(mfrow=c(1, 1))
lines2 <- function(X, Y, type, xlab, ylab, col, pch, lty, lwd, cex) { lines(x=X, y=Y, type="l", col=col) +lines(x=X, y=emaTA(Y), col="black") }
plot(tS3, plot.type="m", col = .colorwheelPalette(3), panel=lines2, grid=TRUE, at="pretty")
```

##Markov chain
Taken from markov chain <https://cran.rstudio.com/web/packages/markovchain/markovchain.pdf>
```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
#create some markov chains
statesNames=c("a","b")
mcA<-new("markovchain", transitionMatrix=matrix(c(0.7,0.3,0.1,0.9),byrow=TRUE,
nrow=2, dimnames=list(statesNames,statesNames)))
plot(mcA)
statesNames=c("a","b","c")
mcB<-new("markovchain", states=statesNames, transitionMatrix=
matrix(c(0.2,0.5,0.3,
0,1,0,
0.1,0.8,0.1),nrow=3, byrow=TRUE, dimnames=list(statesNames,
statesNames)
))
statesNames=c("a","b","c","d")
matrice<-matrix(c(0.25,0.75,0,0,0.4,0.6,0,0,0,0,0.1,0.9,0,0,0.7,0.3),
nrow=4, byrow=TRUE)
mcC<-new("markovchain", states=statesNames, transitionMatrix=matrice)
plot(mcC)
mcD<-new("markovchain", transitionMatrix=matrix(c(0,1,0,1), nrow=2,byrow=TRUE))
plot(mcD)
#operations with S4 methods
mcA^2
steadyStates(mcB)
plot(mcB)
absorbingStates(mcB)
plot(mcB)
markovchainSequence(n=20, markovchain=mcC, include=TRUE)

```
###Various function to perform structural analysis of DTMC Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3,
0, 1, 0,
0.1, 0.8, 0.1), nrow = 3, byrow = TRUE,
dimnames = list(statesNames, statesNames)
))
communicatingClasses(markovB)
recurrentClasses(markovB)
absorbingStates(markovB)
transientStates(markovB)
canonicForm(markovB)
# periodicity analysis : 1
E <- matrix(c(0, 1, 0, 0, 0.5, 0, 0.5, 0, 0, 0.5, 0, 0.5, 0, 0, 1, 0),
nrow = 4, ncol = 4, byrow = TRUE)
mcE <- new("markovchain", states = c("a", "b", "c", "d"),
transitionMatrix = E,
name = "E")
is.irreducible(mcE) #true
period(mcE) #2
# periodicity analysis : 2
myMatr <- matrix(c(0, 0, 1/2, 1/4, 1/4, 0, 0,
0, 0, 1/3, 0, 2/3, 0, 0,
0, 0, 0, 0, 0, 1/3, 2/3,
0, 0, 0, 0, 0, 1/2, 1/2,
0, 0, 0, 0, 0, 3/4, 1/4,
1/2, 1/2, 0, 0, 0, 0, 0,
1/4, 3/4, 0, 0, 0, 0, 0), byrow = TRUE, ncol = 7)
myMc <- new("markovchain", transitionMatrix = myMatr)
period(myMc)
```
###Mobility between income quartiles Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data(blanden)
mobilityMc<-as(blanden, "markovchain")
plot(mobilityMc)
```
###conditionalDistribution of a Markov Chain Example
It extracts the conditional distribution of the subsequent state, given current state.

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
# define a markov chain
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 1, 0, 0.1, 0.8, 0.1),nrow = 3,
byrow = TRUE, dimnames = list(statesNames, statesNames)))
plot(markovB)
conditionalDistribution(markovB, "b")

```

###CD4 cells counts on HIV Infects between zero and six month Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data(craigsendi)
csMc<-as(craigsendi, "markovchain")
steadyStates(csMc)
```

###Function to fit a discrete Markov chain Example
Given a sequence of states arising from a stationary state, it fits the underlying Markov chain distribution using either MLE (also using a Laplacian smoother), bootstrap or by MAP (Bayesian) inference.
```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
sequence <- c("a", "b", "a", "a", "a", "a", "b", "a", "b", "a", "b", "a", "a",
"b", "b", "b", "a")
sequenceMatr <- createSequenceMatrix(sequence, sanitize = FALSE)
mcFitMLE <- markovchainFit(data = sequence)
mcFitBSP <- markovchainFit(data = sequence, method = "bootstrap", nboot = 5, name = "Bootstrap Mc")
mcFitBSP
mcFitMLE
```

###Class "ctmc" Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
energyStates <- c("sigma", "sigma_star")
byRow <- TRUE
gen <- matrix(data = c(-3, 3,
1, -1), nrow = 2,
byrow = byRow, dimnames = list(energyStates, energyStates))
molecularCTMC <- new("ctmc", states = energyStates,
byrow = byRow, generator = gen,
name = "Molecular Transition Model")
steadyStates(molecularCTMC)

```

###Function to fit a CTMC Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data <- list(c("a", "b", "c", "a", "b", "a", "c", "b", "c"), c(0, 0.8, 2.1, 2.4, 4, 5, 5.9, 8.2, 9))
ctmcFit(data)
```

###First passage across states Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
simpleMc <- new("markovchain", states = c("a", "b"),
transitionMatrix = matrix(c(0.4, 0.6, .3, .7),
nrow = 2, byrow = TRUE))
firstPassage(simpleMc, "b", 20)
plot(firstPassage(simpleMc, "b", 20))
```

###Functions to fit a higher order Markov chain Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
library("Rsolnp", lib.loc="~/R/win-library/3.2")
sequence<-c("a", "a", "b", "b", "a", "c", "b", "a", "b", "c", "a", "b", "c", "a"
, "b", "c", "a", "b", "a", "b")
fitHigherOrder(sequence)
```

###Function to obtain the transition matrix from the generator Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
energyStates <- c("sigma", "sigma_star")
byRow <- TRUE
gen <- matrix(data = c(-3, 3,
1, -1), nrow = 2,
byrow = byRow, dimnames = list(energyStates, energyStates))
generatorToTransitionMatrix(gen)
```

###Class "HigherOrderMarkovChain" Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data(holson)
head(holson)
```

###Function to infer the hyperparameters for Bayesian inference from an a priori matrix or a data set Example
Since the Bayesian inference approach implemented in the package is based on conjugate priors, hyperparameters must be provided to model the prior probability distribution of the chain parameters. The hyperparameters are inferred from a given a priori matrix under the assumption that the matrix provided corresponds to the mean (expected) values of the chain parameters. A scaling factor vector must be provided too. Alternatively, the hyperparameters can be inferred from a data set.

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data(rain, package = "markovchain")
inferHyperparam(data = rain$rain)
weatherStates <- c("sunny", "cloudy", "rain")
weatherMatrix <- matrix(data = c(0.7, 0.2, 0.1,
0.3, 0.4, 0.3,
0.2, 0.4, 0.4),
byrow = TRUE, nrow = 3,
dimnames = list(weatherStates, weatherStates))
inferHyperparam(transMatr = weatherMatrix, scale = c(10, 10, 10))
```

###Verify if a state j is reachable from state i. Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames,
transitionMatrix = matrix(c(0.2, 0.5, 0.3,
0, 1, 0,
0.1, 0.8, 0.1), nrow = 3, byrow = TRUE,
dimnames = list(statesNames, statesNames)
)
)
is.accessible(markovB, "a", "c")

```

###Function to check if a Markov chain is irreducible Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b")
mcA <- new("markovchain", transitionMatrix = matrix(c(0.7,0.3,0.1,0.9),
byrow = TRUE, nrow = 2,
dimnames = list(statesNames, statesNames)
))
is.irreducible(mcA)
```

###Class "markovchain" Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
#show markovchain definition
showClass("markovchain")
#create a simple Markov chain
transMatr<-matrix(c(0.4,0.6,.3,.7),nrow=2,byrow=TRUE)
simpleMc<-new("markovchain", states=c("a","b"),
transitionMatrix=transMatr,
name="simpleMc")
#power
simpleMc^4
#some methods
steadyStates(simpleMc)
absorbingStates(simpleMc)
simpleMc[2,1]
t(simpleMc)
is.irreducible(simpleMc)
#conditional distributions
conditionalDistribution(simpleMc, "b")
#example for predict method
sequence<-c("a", "b", "a", "a", "a", "a", "b", "a", "b", "a", "b", "a", "a", "b", "b", "b", "a")
mcFit<-markovchainFit(data=sequence)
predict(mcFit$estimate, newdata="b",n.ahead=3)
#direct conversion
myMc<-as(transMatr, "markovchain")
#example of summary
summary(simpleMc)
## Not run: plot(simpleMc)
plot(simpleMc)
```

###Class "markovchainList" Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
showClass("markovchainList")
#define a markovchainList
statesNames=c("a","b")
mcA<-new("markovchain",name="MCA", transitionMatrix=matrix(c(0.7,0.3,0.1,0.9),
byrow=TRUE, nrow=2, dimnames=list(statesNames,statesNames)))
mcB<-new("markovchain", states=c("a","b","c"), name="MCB",
transitionMatrix=matrix(c(0.2,0.5,0.3,0,1,0,0.1,0.8,0.1),
nrow=3, byrow=TRUE))
mcC<-new("markovchain", states=c("a","b","c","d"), name="MCC",
transitionMatrix=matrix(c(0.25,0.75,0,0,0.4,0.6,
0,0,0,0,0.1,0.9,0,0,0.7,0.3), nrow=4, byrow=TRUE)
)
mcList<-new("markovchainList",markovchains=list(mcA, mcB, mcC),
name="Non - homogeneous Markov Chain")
```

###markovchainListFit Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
# using holson dataset
data(holson)
# fitting a single markovchain
singleMc <- markovchainFit(data = holson[,2:12])
# fitting a markovchainList
mclistFit <- markovchainListFit(data = holson[, 2:12], name = "holsonMcList")
```

###Function to generate a sequence of states from homogeneous Markov chains. Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
# define the markovchain object
statesNames <- c("a", "b", "c")
mcB <- new("markovchain", states = statesNames,
transitionMatrix = matrix(c(0.2, 0.5, 0.3, 0, 0.2, 0.8, 0.1, 0.8, 0.1),
nrow = 3, byrow = TRUE, dimnames = list(statesNames, statesNames)))
# show the sequence
outs <- markovchainSequence(n = 100, markovchain = mcB, t0 = "a")
```

###A function to compute multinomial confidence intervals of DTMC Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
seq<-c("a", "b", "a", "a", "a", "a", "b", "a", "b", "a", "b", "a", "a", "b", "b", "b", "a")
mcfit<-markovchainFit(data=seq,byrow=TRUE)
seqmat<-createSequenceMatrix(seq)
multinomialConfidenceIntervals(mcfit$estimate@transitionMatrix, seqmat, 0.95)
```

###Method to retrieve name of markovchain object Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 1, 0, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames=list(statesNames,statesNames)),
name = "A markovchain Object"
)
name(markovB)

statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 1, 0, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames=list(statesNames,statesNames)),
name = "A markovchain Object"
)
name(markovB) <- "dangerous mc"

```

###Function to compute the probability of observing a new data set, given a data set Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
sequence<-c("a", "b", "a", "a", "a", "a", "b", "a", "b", "a", "b", "a", "a",
"b", "b", "b", "a")
hyperMatrix<-matrix(c(1, 2, 1, 4), nrow = 2,dimnames=list(c("a","b"),c("a","b")))
predProb <- predictiveDistribution(sequence[1:10], sequence[11:17], hyperparam =hyperMatrix )
hyperMatrix2<-hyperMatrix[c(2,1),c(2,1)]
predProb2 <- predictiveDistribution(sequence[1:10], sequence[11:17], hyperparam =hyperMatrix2 )
predProb2==predProb
```

###Preprogluccacon DNA protein bases sequences Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data(preproglucacon)
preproglucaconMc<-markovchainFit(data=preproglucacon$preproglucacon)
```

###Prior Distribution Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
priorDistribution(matrix(c(0.5, 0.5, 0.5, 0.5),
nrow = 2,
dimnames = list(c("a", "b"), c("a", "b"))),
matrix(c(2, 2, 2, 2),
nrow = 2,
dimnames = list(c("a", "b"), c("a", "b"))))
```

###Alofi island daily rainfall Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data(rain)
rainMc<-markovchainFit(data=rain$rain)

```

###Function to generate a sequence of random CTMC transitions Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
energyStates <- c("sigma", "sigma_star")
byRow <- TRUE
gen <- matrix(data = c(-3, 3,
1, -1), nrow = 2,
byrow = byRow, dimnames = list(energyStates, energyStates))
molecularCTMC <- new("ctmc", states = energyStates,
byrow = byRow, generator = gen,
name = "Molecular Transition Model")
statesDist <- c(0.8, 0.2)
rctmc(n = Inf, ctmc = molecularCTMC, T = 1)
rctmc(n = 5, ctmc = molecularCTMC, initDist = statesDist, include.T0 = FALSE)
```

###Function to generate a sequence of states from homogeneous or nonhomogeneous Markov chains Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
# define the markovchain object
statesNames <- c("a", "b", "c")
mcB <- new("markovchain", states = statesNames,
transitionMatrix = matrix(c(0.2, 0.5, 0.3, 0, 0.2, 0.8, 0.1, 0.8, 0.1),
nrow = 3, byrow = TRUE, dimnames = list(statesNames, statesNames)))
# show the sequence
outs <- rmarkovchain(n = 100, object = mcB, what = "list")
#define markovchainList object
statesName = c("a", "b", "c")
mcA <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 0.2, 0.8, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames = list(statesNames, statesNames)))
mcB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 0.2, 0.8, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames = list(statesNames, statesNames)))
mcC <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 0.2, 0.8, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames = list(statesNames, statesNames)))
mclist <- new("markovchainList", markovchains = list(mcA, mcB, mcC))
# show the list of sequence
rmarkovchain(100, mclist, "list")
```

###Sales Demand Sequences Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
data("sales")
#fitHighOrderMultivarMC(seqMat = sales, order = 2, Norm = 2)
```

###Defined states of a transition matrix Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 1, 0, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames=list(statesNames,statesNames)),
name = "A markovchain Object"
)
states(markovB)
```

###SStationary states of a markovchain objeect Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 1, 0, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames=list(statesNames,statesNames)),
name = "A markovchain Object"
)
steadyStates(markovB)

```

###Function to get the transition probabilities from initial to subsequent states Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
statesNames <- c("a", "b", "c")
markovB <- new("markovchain", states = statesNames, transitionMatrix =
matrix(c(0.2, 0.5, 0.3, 0, 1, 0, 0.1, 0.8, 0.1), nrow = 3,
byrow = TRUE, dimnames=list(statesNames,statesNames)),
name = "A markovchain Object"
)
transitionProbability(markovB,"b", "c")
```

###Various functions to perform statistical inference of DTMC Example

```{r}
library("markovchain", lib.loc="~/R/win-library/3.2")
sequence <- c("a", "b", "a", "a", "a", "a", "b", "a", "b",
"a", "b", "a", "a", "b", "b", "b", "a")
mcFit <- markovchainFit(data = sequence, byrow = FALSE)
verifyMarkovProperty(sequence)
assessOrder(sequence)
assessStationarity(sequence, 1)
divergenceTest(sequence, mcFit$estimate@transitionMatrix)
```

###read from csv Example

```{r}
library("csvread", lib.loc="~/R/win-library/3.2")
cic = csvread("C:\\Users\\ricca\\Documents\\R\\intermmidiate-ud.csv", coltypes = c("string","double","double","integer","integer","integer","long","string","integer","string"),header = TRUE)
summary(cic)
```

###read from csv read csv Example

```{r}
library("csvread", lib.loc="~/R/win-library/3.2")
cic = read.csv("C:\\Users\\ricca\\Documents\\R\\intermmidiate-ud.csv",header = TRUE, nrows = 10000)
summary(cic)
plot(cic$sys,cic$modification)
```

###ddply example group
```{r}
library("plyr", lib.loc="~/R/win-library/3.2")
.(a, b, c)
.(first = a, second = b, third = c)
.(a ^ 2, b - d, log(c))
as.quoted(~ a + b + c)
as.quoted(a ~ b + c)
as.quoted(c("a", "b", "c"))
# Some examples using ddply - look at the column names
ddply(mtcars, "cyl", each(nrow, ncol))
ddply(mtcars, ~ cyl, each(nrow, ncol))
ddply(mtcars, .(cyl), each(nrow, ncol))
ddply(mtcars, .(log(cyl)), each(nrow, ncol))
ddply(mtcars, .(logcyl = log(cyl)), each(nrow, ncol))
ddply(mtcars, .(vs + am), each(nrow, ncol))
ddply(mtcars, .(vsam = vs + am), each(nrow, ncol))

```

###ggplot example group
<http://stackoverflow.com/questions/16961921/plot-data-in-descending-order-as-appears-in-data-frame>
```{r}
set.seed(42)
df <- data.frame(Category = sample(LETTERS), Count = rpois(26, 6))
require("ggplot2")
p1 <- ggplot(df, aes(x = Category, y = Count)) +
         geom_bar(stat = "identity")
p2 <- ggplot(df, aes(x = reorder(Category, -Count), y = Count)) +
         geom_bar(stat = "identity")
require("gridExtra")
grid.arrange(arrangeGrob(p1, p2))
```

###ggplot example group
```{r}
library("csvread", lib.loc="~/R/win-library/3.2")
cic = read.csv("C:\\Users\\ricca\\Documents\\R\\intermmidiate-ud.csv",header = TRUE, nrows = 10000)
summary(cic)

require("ggplot2")
p1 <- ggplot(cic, aes(x = system, y = modification)) +
         geom_bar(stat = "identity") + theme(axis.text.x=element_text(angle = 90, hjust = 0.5))
p2 <- ggplot(cic, aes(x = reorder(system, Commit), y = modification)) +
         geom_bar(stat = "identity") + theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
require("gridExtra")
grid.arrange(arrangeGrob(p1, p2))
```